---
title: "China's One Child Policy"
author: "Yingtong Guo"
date: "2020/2/12"
output: html_document
---

``````{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(message=FALSE)
knitr::opts_chunk$set(warning=FALSE)
```

```{r viz1}
library(tidyverse)
library(ggplot2)
library(readxl)
library(ggthemes)
data <- read_excel("data/data.xlsx")
high<-c("China","India","Philipines","South Korea","Thailand")
data2 <- filter(data, country %in% high) %>%
  mutate(country2=country)
data2 %>%
  ggplot(aes(x=year, y=fertility_rate)) +
  geom_line(aes(color=country), color="royalblue4", size=1, alpha=0.5 )+
  geom_point(aes(color=country), color="royalblue4", size=1.2 )+
  geom_vline(xintercept=c(1978, 2013), linetype="dotted")+
  theme(legend.position="none")+
  theme_pander(base_size = 9, base_family = "sans")+
  scale_y_continuous(limits=c(0, 8))+
  labs(x="",
       y="Fetility rate, total (births per woman)",
       title="Total Fertility Rate of China, India, Philipines, South Korea, and Thailand (1959-2017)",
       subtitle=paste("All these countries in Asia experienced rapid declines in fertility during economic development"),
       caption="Data Source: World Bank")+
  facet_wrap(~country)
```

China implemented a strict one-child policy between 1978 and 2013, and then published a loosened policy allowing parents to have two kids in total. People argue that the policy is the main reason why China's fertility rate decreased dramatically. However, when we compare China to other Asian countries with similar fertility rates in the 1960s, these countries not restricted by the one-child policy still exhibited rapid declines in fertility rates.